---
title: "p8131_hw3_jsg2145"
author: "Jared Garfinkel"
date: "2/17/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(ResourceSelection)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 1

```{r}
case_df = tibble(
  age = seq(from = 25, to = 75, by = 10),
  unexposed = c(0, 5, 21, 34, 36, 8),
  exposed = c(1, 4, 25, 42, 19, 5)
) %>% 
  mutate(case = 1)

control_df = tibble(
  age = seq(from = 25, to = 75, by = 10),
  unexposed = c(106, 164, 138, 139, 88, 31),
  exposed = c(9, 26, 29, 27, 18, 0)
) %>% 
  mutate(case = 0)

df = union(case_df, control_df) %>% 
  pivot_longer(cols = c(exposed, unexposed), names_to = "status") %>% 
  mutate(status = recode(status,
                         exposed = 1,
                         unexposed = 0))
```

```{r}
fit = glm(cbind(value*(case == 1), value*(case == 0)) ~ status + age, family = binomial(link = 'logit'), data = df)

summary(fit)

fit = fit %>% 
  broom::tidy()
```

The difference in log odds of developing cancer is `r fit %>% filter(pull(fit, term) == "status") %>% pull(estimate) %>% round(digits = 2)` for those with the exposure (>80 grams of alcohol per day) compared to those without, holding age constant.

The log odds ratio of developing cancer is `r fit %>% filter(pull(fit, term) == "age") %>% pull(estimate) %>% round(digits = 2)` for each 1 unit increase in age, holding exposure constant.

## Problem 2

```{r}
bean_75 = tibble(
  y = c(10, 23, 23, 26, 17),
  m = c(39, 62, 81, 51, 39)) %>% 
  mutate(extract = "bean",
         seed = "75")

cuc_75 = tibble(
  y = c(5, 53, 55, 32, 46, 10),
  m = c(6, 74, 72, 51, 79, 13)) %>% 
  mutate(extract = "cuc",
         seed = "75")

bean_73 = tibble(
  y = c(8, 10, 8, 23, 0),
  m = c(16, 30, 28, 45, 4)) %>% 
  mutate(extract = "bean",
         seed = "73")

cuc_73 = tibble(
  y = c(3, 22, 15, 32, 3),
  m = c(12, 41, 30, 51, 7)) %>% 
  mutate(extract = "cuc",
         seed = "73")

df_orobanche = union_all(bean_75, cuc_75) %>% 
  union_all(bean_73) %>% 
  union_all(cuc_73) %>% 
  mutate(y_m = m - y,
         extract = recode(extract,
                          "bean" = 0,
                          "cuc" = 1),
         seed = recode(seed,
                       "75" = 0,
                       "73" = 1))

fit_o = glm(cbind(y, y_m) ~ seed + extract, family = binomial(link = 'logit'), df_orobanche)

summary(fit_o)
```

The log odds ratio for germination is reduced by `r broom::tidy(fit_o) %>% filter(term == "seed") %>% pull(estimate) %>% round(digits = 2)` for species Orobanche aegyptiaca 73 compared to Orobanche aegyptiaca 75, holding the extract media type constant.

The log odds ratio for germination is increased by `r broom::tidy(fit_o) %>% filter(term == "extract") %>% pull(estimate) %>% round(digits = 2)` for cucumber extract media compared to bean extract media, holding the species constant.

There is a log odds of `r broom::tidy(fit_o) %>% filter(term == "(Intercept)") %>% pull(estimate) %>% round(digits = 2)` for plants that are species Orobanche aegyptiaca 75 grown in bean extract media.

```{r}
# Evaluate dispersion with Pearson Chi-Squared Residual

dev_o = deviance(fit_o)

pval_x2 = 1 - pchisq(dev_o, 19)
pval_x2
```

A low p-value (<0.05) indicates that there is over dispersion

```{r}
# Estimate phi with the G statistic

G.stat = sum(residuals(fit_o, type = "pearson")^2)
G.stat

phi = G.stat/(21 - 2)
phi

# estimate with the deviance

dev_o/(21 - 2)
```

```{r}
res_o = residuals(fit_o, type = "pearson")
plot(qnorm((21+1:21+0.5)/(2*21+1.125)),sort(abs(res_o)),xlab='Expected Half-Normal Order Stats',ylab='Ordered Abs Pearson Residuals')
abline(a=0,b=1)
abline(a=0,b=sqrt(phi),lty=2)
```

```{r}
summary(fit_o, dispersion=phi)
```

The type of seed is no longer significant in the model.

The reason for the over dispersion may be correlation within each subject and treatment group.

